Towards a framework for reasoning about stateful compiler IRs using coinductive reasoning.

### 2024-06-25 paper.tex
fix timeline

CompCert is the
seminal work when it comes to verified compilation: an end-to-end
verified to not miscompile compiler for the C language~\cite{leroyCompCertFormallyVerified}.
However, exactly
because it's an end-to-end verified monolith, CompCert is not
competitive compared to traditional non-verified compilers. Outside of
certain high-risk domains, where traditionally optimizations would be
turned off entirely,~\cite{kastnerCompCertPracticalExperience} the
vast majority of C code is thus still compiled with non-verified
compilers, which studies have shown to be prone to miscompilations~\cite{yangFindingUnderstandingBugs2011}.

@@

Our group has been developing
LeanMLIR, a framework for giving
semantics to MLIR-style compiler IRs in the Lean theorem prover~\cite{demouraLeanTheoremProver2015}. 
As a theorem prover, Lean is
much more expressible than SMT solvers. Among theorem provers, Lean is
particularly suited for this purpose because of it's community of
mathematicians developing a comprehensive mathematical library,~\cite{themathlibcommunityLeanMathematicalLibrary2020} 
and because of its highly extensive syntax and meta-programming capabilities~\cite{ullrichNotationsHygienicMacro2022, paulinoMetaprogrammingLean}.

@@

\subsubsection{Runtime behaviour}\label{runtime-behaviour}
\begin{quote}
September '25 - May '25
\end{quote}

>>>>

CompCert is the
seminal work when it comes to verified compilation: an end-to-end
verified to not miscompile compiler for the C language~\cite{leroyCompCertFormallyVerified}.
However, exactly
because it's an end-to-end verified monolith, CompCert is not
competitive compared to traditional non-verified compilers. Outside of
certain high-risk domains, where traditionally optimizations would be
turned off entirely~\cite{kastnerCompCertPracticalExperience},
the vast majority of C code is thus still compiled with non-verified
compilers, which studies have shown to be prone to miscompilations~\cite{yangFindingUnderstandingBugs2011}.

@@

Our group has been developing
LeanMLIR, a framework for giving
semantics to MLIR-style compiler IRs in the Lean theorem prover~\cite{demouraLeanTheoremProver2015}. 
As a theorem prover, Lean is
much more expressible than SMT solvers. Among theorem provers, Lean is
particularly suited for this purpose because of it's community of
mathematicians developing a comprehensive mathematical library~\cite{themathlibcommunityLeanMathematicalLibrary2020},
and because of its highly extensive syntax and meta-programming capabilities~\cite{ullrichNotationsHygienicMacro2022, paulinoMetaprogrammingLean}.

@@

\subsubsection{Runtime behaviour}\label{runtime-behaviour}
\begin{quote}
September '25 - Jan '26
\end{quote}
